\Large \bf Cyclone: Replication Middleware for NVM Applications

### 2016-09-22 paper/paper.tex
Added cc results.
Added app results.
Edits.

\documentclass[letterpaper,twocolumn,10pt]{article}
\usepackage{usenix,epsfig,endnotes,graphicx}
\begin{document}

@@

% Use the following at camera-ready time to suppress page numbers.
% Comment it out when you first submit the paper for review.
\thispagestyle{empty}

\subsection*{Abstract}
Memory technology in the datacenter is due to undergo a paradigm shift with the
increased use of directly addressable non-volatile memory both in the form of
battery backed non-volatile DIMMs as well as newer memory types such as
3D XPoint. In anticipation, libraries that allow programmers to build durable
data structures on non-volatile heaps have become available. A missing piece
however is a component to make such durable data structures available using
replication across a commodity network such as ethernet. Existing solutions that
offer good performance are biased towards pure transactions across RDMA with a
dependence on external services and application specific code for fault recovery.

Cyclone is replication middleware that aims to add fault tolerance as
transparently as possible. A design for Cyclone is that NVM application
developers should not need to add fault recovery code. A second design goal for
Cyclone is to strive for good performance on commodity ethernet - with
innovations in the software stack rather than a requirement for RDMA capable
networking hardware.

Cyclone achieves these goals by providing a simple API for client server
applications that allows server state to be replicated by executing the same
sequence of RPC calls across all replicas. It provides at most once semantics
with automatic failover to hide fault recovery from programmers. Finally it uses
DPDK - a high performance, low latency network stack coupled with batching,
zero-copy and the ability to scale replication performance across
multiple instances of a consensus protocol to replicate millions of small
payload RPC messages a second on commodity ethernet.

\section{Introduction}
TBD
\section{Single Node Programming Model}
Cyclone starts with the assumption that the programmer has already built a
concurrent durable application on a non-volatile heap. A number of libraries and
filesystems make this possible today [DAX, NVML, others ?]. As a starting point,
we support applications built on top of Intel's Non Volatile Memory Library
(NVML) [nvml.io]. We describe NVML's programming model in this section to
provide adequate background for the rest of this paper. The primary feature of
interest is the ability to write code that is crash consistent. In other words
code that makes changes to the non-volatile heap is enclosed in a crash
consistent transaction such that either all of the transaction executes or none
of it executes in the face of power failures. This is critical to maintaining
consistency as the data structure, being durable, must survive across power
failures. To illustrate this idea, Figure~\ref{fig:example} shows how one might
build a persistent linked list using NVML - the key idea is to enclose the
entire process of updating the linked list within a crash consistent
transaction. This ensure that the linked list is not corrupted due to a power
failure. A second point worth noting is that NVM libraries such as NVML usually
provide a way to wrap pointers to non-volatile memory (using {\tt TOID} macros)
in order that such memory may be mapped to different locations in virtual memory
across restarts without impact to the accessing application.

@@

It is easy to turn the linked list example in NVML into a concurrent linked list
by simply adding a lock to it. Such a lock can be maintined in volatile DRAM if
so desired for performance. More sophisticated fine-grained or optimistic
concurrency control mechanisms for NVM are also possible [mnemosyne].

\section{Replication}
Cyclone is designed to add fault tolerance to NVML client server applications
via replication - a quorum of machines maintains \emph{equivalent} state. Server
state is queried and manipulated via RPC calls from clients. Cyclone replicates
the RPC call itself across replicas rather than replicating every access made
during execution of the RPC call.

Cyclone provides strongly consistent replication - every machine in the quorum
of replicas maintains a log of RPC calls. Cyclone treats the RPC call itself as
a variable sized opaque blob of data, leaving the application free to choose how
it wishes to marshall call related details and argument. All machines agree on
the committed sequence of RPC calls in the log. We achieve this by running an
instance of the RAFT [raft] consensus protocol to keep the log of RPC calls on
different machines in synchronization. Specially flagged `read-only` RPC calls
are executed on the leader replica and are not replicated.

\subsection{Decoupled Replication}
\label{sec:decouple}
The standard replicated state machine approach requires that commands (entries
in the log) be committed before they can be applied to the state machine i.e. in
our case - the RPC call be executed. In contrast, Cyclone decouples execution of
the RPC call on any replica from its replication.

However, the fact that RPC calls that modify state must themselves be executed
in a crash consistent transaction allows us to overlap execution with
replication. Normally, this would be a problem as replication might failt for a
number of reasons but primarily due to a change of leader and subsequent of
rollback of previously appended (but not committed) log entries. To solve this
problem, Cyclone delays the commit of the failure atomic transaction after
execution till replication is complete. In the event that replication fails, the
transaction is aborted and any changes to the non volatile heap are rolled back.

A pseudocode description of how the runtime handles \emph{decoupled replication}
is shown in Figure~\ref{fig:async_rep}.

@@

Asynchronous replication provides good performance and we believe it would be the
common case for RPC calls replicated via Cyclone. However the program might on
occasion require successful replication to be a precondition to exection. A case
where we find this necessary is output commit - where the RPC call has external
side effects, such as communicating with other machines in the cluster and
therefore crash consistency is insufficient to roll back the execution of the
RPC call in the event that replication fails. We therefore provide users the
ability to declare at the client side that an RPC call should execute
synchronously at the server side. This results in Cyclone ensuring that the RPC
call is successfully replicated before it attempts execution on any of the
replicas.

\subsection{Network Stack}
Cyclone is designed around DPDK [dpdk] a high performance network stack. DPDK
provides convenient access to packets arriving at an ethernet NIC by moving them
into packet buffers with the contained cache lines also directly placed into the
last level cache of the CPU during DMA [ddio].

Cyclone separates the available CPU cores on a machine, in a configurable way,
into cores dedicated to work related to the consensus protocol and cores
decidated to running application code. Each RAFT core runs an \emph{independent}
instance of the RAFT protocol. RAFT cores pass work to the application cores
through lock-free double ended queues. A RAFT core dispatches an incoming RPC
call for execution simultaneously with sending it out for replication.

Each RAFT core has a dedicated NIC queue pair. The input queue receives RPC
calls from clients as well as RAFT protocol related traffic. The output queue
only carries RAFT protocol related traffic. Each application core has a
dedicated \emph{output} queue on the NIC. When completed, the response to the
RPC call is sent from the application core via its dedicated output queue. This
division of NIC queues eliminates any need for synchronization between RAFT
cores and application cores when accessing the NIC.

In the event that the machine has multiple NICs Cyclone binds each RAFT or
application core to a specific NIC and picks the necessary queues from that
NIC. Spreading the network demands from the cores across the NICs in this manner
allows Cyclone to scale with multiple NICs when available.

@@

In order to avoid copying the incoming RPC call contents to a persistent memory
range, Cyclone uses persistent memory on the system itself to allocate space for
packet buffers. However although a packet is accessible at the core the most up
to date data for the memory region occupied by the packet is in the CPU cache
and not necessarily in memory. To ensure that packet data has truly been made
persistent, Cyclone issues a series of cacheline flushes. This ensures that the
packet data is persistent. Cyclone maintains its log as a circular sequence of
pointers to packet buffers and therefore the append operation is completed by
adding a pointer to the received packet and flushing the cacheline containing
the pointer.

We exploit the fact that DPDKs packet buffers allow prepending data to a packet
without physically moving the data in the packet buffer. Fig
[figures/packets.bmp] shows how the layout of a packet buffer changes at the
leader during replication. The received packets contains only the RPC call from
the client. We prepend a RAFT related header (containing the previous log index,
previous log term, leader's term and leader's commit point) to packet buffer
itself. To this RAFT header we then prepend an IPv4 header. We set IP addresses
in the header to use 5-tuple steering in order to place the packet in the right
queue at the receiving NIC. We do not use different ip addresses for different
replicas, rather the ip addresses are only set in conjunction with NIC
programming to enable steering of the received packets to the correct NIC queue.
This composite packet buffer is then sent \emph{unchanged} to every replica. In
order to properly route the packet through switches, we allocate a small buffer
containing an ethernet header for each follower replica and \emph{chain} it in
front of the prepared packet - ensuring we do not need to change the contents of
the packet itself for every replica.

Although the common case path in RAFT is straightforward, it still adds many
hundreds of cycles for each log entry to be processed. This adds up to a few
microseconds, a significant amount in the face of sub 10 us network latency
between nodes with DPDK. We amortize this overhead when operating under load by
batching multiple RPC calls together and treating the batch as a single RAFT log
entry.  DPDK by itself can efficiently batch incoming packets under load -
allowing the user to easily receive a burst (currently limited to 32) of packets
rather than a single one for a polling call. DPDK further allows chaining packet
buffers together and treating the chain as a single packet. This is convenient
for our zero copy approach and we chain packet mbufs together to form a single
packet as shown in Fig [figures/chain.bmp]. Since the entire chain is sent as a
single packet when being replicated, the length of a chain is limited by the
size of the coalesced packet, that must fit into an ethernet jumbo frame of
approximately 9KB.

@@

\subsection{Distributed Transactions}
A useful primitive for many distributed systems is the ability to run
distributed transactions - an atomic sequence of conditional updates spanning
multiple machines. Distributed transactions have two distinct facets: atomicity
and fault tolerance. Distributed transactions appear atomic with respect to
other atomic transactions and simple updates or reads by virtue of standard
concurrency control techniques. A common technique is strict two phase
locking: acquire all locks, update the locked objects and finally release all
the locks. We believe that NVM application programmers - particularly those
having experience with concurrent code should not find it hard to deal with
distributed concurrency. The hard part for most programmers is likely to be the
second aspect: fault tolerance. A distributed transaction spans many machines
and has a large footprint to clean up in the event of failure. Failure handling
is made more complex when considering cascading failures - further failures
while recovering from a failure. Cyclone was designed to allow primitives such
as distributed transctions to be added without the programmer needing to worry
about faults or write fault recovery code. As an exercise we show how to add
support for distributed transactions to our hash table.

@@

We guard against faults by executing the RPC call on a quorum of
machines via Cyclone. Since execution of the RPC call leaves \emph{externally
  visible side effects} by virtue of the RPC calls it makes we require output
commit [] - the property that the execution is guaranteed to complete once it
starts. This is simply achieved by a flag in the RPC call that turns off
decoupled execution (Section~\ref{sec:decouple}). The flag causes execution to
start only after the RPC call has been successfully replicated.

@@

We note that this design is not intended for performance, under the assumption
that distributed transactions will be infrequent compared to single key
operations.  On the other hand, it is worth comparing the lack of any need for
explicit fault recovery in Cyclone with fault recovery in systems systems that
explicity require such cleanup [FARM, also refer to pages].  We believe that
such designs, which make availability easy for NVM applications are key to
increasing the adoption rate of NVM by data center operators.

@@

\section{Evaluation}

\begin{figure*}
\begin{tabular}{cc}
\begin{minipage}{0.5\textwidth}
  \includegraphics[width=\textwidth,height=6cm]{results/replication_mops_0.pdf}
\end{minipage}&
\begin{minipage}{0.5\textwidth}
  \includegraphics[width=\textwidth,height=6cm]{results/replication_gbps_0.pdf}
\end{minipage}\\
\begin{minipage}{0.5\textwidth}
  \includegraphics[width=\textwidth,height=6cm]{results/replication_mops_512.pdf}
\end{minipage}&
\begin{minipage}{0.5\textwidth}
  \includegraphics[width=\textwidth,height=6cm]{results/replication_gbps_512.pdf}
\end{minipage}
\end{tabular}
\caption{Pure Replication}
\label{fig:pure_rep}
\end{figure*}

\section{Conclusion}

>>>>

\documentclass[letterpaper,twocolumn,10pt]{article}
\usepackage{usenix,epsfig,endnotes,graphicx,url}
\begin{document}

@@

% Use the following at camera-ready time to suppress page numbers.
% Comment it out when you first submit the paper for review.
%\thispagestyle{empty}

\subsection*{Abstract}
The introduction of non-volatile directly attached memory in datacenter - in
the form of NVDIMMs and new memory technologies allows efficient data structures
to be built on non-volatile heaps that provide good performance coupled with
durability. However high availability remains a problem for these primarily
single machine solutions. While transactional interfaces built over RDMA
capable networking hardware offer a viable alternative, they are often
unavailable in datacenters. Cyclone solves this problem by providing transparent
replication middleware that to be used with Intel's Non Volatile Memory Library 
(NVML) to construct durable data-structures that map keys to values.
Cyclone is designed around an optimized userspace networking stack to reduce
latency and can scale its consensus protocol across cores and network ports to
provide good throughput under load. We show that Cyclone can replicate as many
as 8 million updates a second using just 4 commodity 10 GigE ports on a machine.

\section{Introduction}
Memory technology in datacenters is due to undergo a paradigm shift with the
increased use of directly addressable non-volatile memory both in the form of
battery backed non-volatile DIMMs~\cite{farm} as well as newer memory types
such as 3D XPoint~\cite{pmfs, bpfs}. In anticipation, libraries that allow
programmers to build durable data structures on non-volatile heaps have become
available~\cite{mnemosyne, nvheaps, cdds}  - an example of a such a
library that we have used in this paper is Intel's non volatile memory library
(NVML)~\cite{nvml}. A missing piece however is a component to make such durable 
data structures highly available using replication across a commodity network
such as ethernet. Existing solutions that offer good performance are biased
towards pure transactions across RDMA with a dependence on external services
(such as zookeeper~\cite{zookeeper}) and application specific code for
fault recovery~\cite{farm, htm}.
This is a problem for programmers using NVML to construct durable and available 
data structures where one cannot apriori guarantee the availability of
RDMA (for e.g., when running on externally hosted environments or public
clouds).

Cyclone is a prototype replication middleware that aims to demonstrate that one
can add fault tolerance transparently to a specific class of client server
applications written on top of libraries such as NVML. We specifically target
key-value maps in this paper and in that context make two key
contributions. First, we propose a simple API that enables server state to be
replicated without the programmer needing to worry about fault recovery. This
allows one to start by building a concurrent durable data structure using
libraries such as NVML and add availability at a later stage using Cyclone. The
second contribution is to show that one can achieve reasonably good replication 
performance using commodity ethernet via innovations in the networking stack and
by scaling replication with the application. This result should be a useful
design point for developers looking to deploy applications on non-volatile heaps
in commodity ethernet-based environments such as public clouds.

\section{Single Node Programming Model}
Cyclone assumes a concurrent durable application on non-volatile heap as the
starting point. A number of libraries and filesystems make this possible
today~\cite{nvml, dax, pmfs, mnemosyne, nvheaps, cdds}. Cyclone currently
supports applications built on top of Intel's Non Volatile Memory Library
(NVML~\cite{nvml}). We describe NVML's programming model in this section to
provide adequate background for the rest of this paper. NVML requires 
that code that makes changes to non-volatile heap is enclosed in a crash
consistent transaction such that either all of the transaction executes or none
of it executes in the event of power failures.
%This is critical to maintaining consistency as the data structure, being
%durable, must survive across power failures.
Figure~\ref{fig:example} illustrates the use of NVML to build a persistent
linked list. Updates to the linked list are enclosed in a NVML transaction to
avoid corruption due to power failure. It is also worth noting that NVM
libraries such as NVML usually allow remapping non-volatile heaps to a different
virtual memory address range across restarts (e.g., using {\tt TOID} macros to
implement fat pointers in NVML).

@@

It is easy to turn the linked list example in NVML into a concurrent linked list
by simply adding a lock to it. Such a lock can be maintined in volatile DRAM if
so desired for performance. More sophisticated fine-grained or optimistic
concurrency control mechanisms for NVM are also
possible~\cite{mnemosyne, cdds, nvheaps}.

\section{Replication}
Cyclone is designed to add fault tolerance to NVML client server applications
via replication -- a quorum of machines maintains \emph{equivalent} state.
Server state is queried and manipulated via RPC calls from clients.
Cyclone replicates the RPC call itself across replicas rather than replicating
every access (to the server state) made during execution of the RPC call.

Cyclone provides strongly consistent replication - every machine in the quorum
of replicas maintains a log of RPC calls. Cyclone treats the RPC call itself as
a variable sized opaque blob of data, leaving the application free to choose how
it wishes to marshall call related details and arguments. All machines agree on
the committed sequence of RPC calls in the log. We achieve this by running an
instance of the RAFT~\cite{raft} consensus protocol to keep the log of RPC calls
on different machines in synchronization. Specially flagged `read-only` RPC
calls are executed only on the leader replica and are not replicated.

\subsection{Decoupled Replication}
\label{sec:decouple}
The standard replicated state machine approach requires that commands (entries
in the log) be committed before they can be applied to the state machine --
i.e. in our case before the RPC call can be executed. In contrast, Cyclone
decouples execution of an RPC call on any replica from its replication.
The fact that RPC calls that modify state must themselves be executed
in a crash consistent transaction allows us to overlap execution with
replication. Normally, this would be a problem as replication might fail for a
number of reasons but primarily due to change of leadership and subsequent
rollback of previously appended (but not committed) log entries. To solve this
problem, Cyclone delays the actual commit of the failure atomic transaction
until after the replication is complete. In the event that replication fails,
the transaction is aborted and any changes to the non volatile heap are rolled
back. Figure~\ref{fig:async_rep} shows a pseudocode description of how the
runtime handles \emph{decoupled replication}.

@@

Asynchronous replication provides good performance and we believe it would be
the common case for RPC calls replicated via Cyclone. However the program might 
on occasion require successful replication to be a precondition to exection.
A case where we find this necessary is
\textit{output commit}~\cite{output_commit} - where the RPC call has
external side effects (e.g., communication with other machines in the cluster)
and therefore crash consistency is insufficient to roll back the execution of
the RPC call in the event that replication fails. We therefore provide users the
ability to declare at the client side that an RPC call should execute
synchronously at the server side. Cyclone would then ensure that the RPC
call is successfully replicated before it attempts execution on any of the
replicas.

\subsection{Network Stack}
Cyclone is designed around DPDK~\cite{dpdk} -- a high performance network stack
optimized for Intel processors. DPDK provides convenient access to packets
arriving at an ethernet NIC by moving them into packet buffers and directly
placing those buffers into the last level cache of the CPU during DMA~\cite{ddio}.

Cyclone separates the available CPU cores on a machine, in a configurable way,
into cores dedicated to work related to the (RAFT) consensus protocol and cores
decidated to running application code. Each RAFT core runs an \emph{independent}
instance of the RAFT protocol. RAFT cores pass work to the application cores
through lock-free double ended queues. A RAFT core dispatches an incoming RPC
call for execution simultaneously with sending it out for replication.

Each RAFT core has a dedicated NIC queue pair. The input queue receives RPC
calls from clients as well as RAFT protocol related traffic. The output queue
only carries RAFT protocol related traffic. Each application core has a
dedicated \emph{output} queue on the NIC. When completed, the response to the
RPC call is sent from the application core via its dedicated output queue. This
division of NIC queues eliminates the need for synchronization between RAFT
cores and application cores when accessing the NIC.

In the event that the machine has multiple NICs, Cyclone spreads the network
demands and scales by binding each RAFT or application core to a specific NIC
and picking the necessary queues from that NIC.

@@

In order to avoid an additional copy of the incoming RPC call contents from DRAM
to persistent memory, Cyclone allocates space for packet buffers on non-volatile
heap. However, since packet data is cached, Cyclone ensures durability
(or persistence) of this data by explicitly flushing the dirty
cachelines~\cite{pmfs}. Cyclone maintains its log as a circular sequence of
pointers to packet buffers and therefore the append operation is completed by
adding a pointer to the received packet and flushing the cacheline containing
the pointer.

We exploit the fact that DPDK allows prepending data to a packet
without physically moving the data in a packet buffer. The received packet
contains only the RPC call from the client. We prepend a RAFT related header
(containing the previous log index, previous log term, leader's term and
leader's commit point~\cite{raft}) to packet buffer itself. To this RAFT header 
we then prepend an IPv4 header. We assign pre-determined values to the
IP address fields in order to exploit 5-tuple steering at the receiving NIC for 
the placement of the incoming packet in the right queue.
This composite packet buffer is then sent \emph{unchanged} to every replica.
To properly route the packet through switches, we allocate a small buffer
containing an ethernet header for each follower replica and \emph{chain} the
buffer in front of the prepared packet - thereby ensuring that we do not need
to change the contents of the packet itself for every replica.

Although the common case path in RAFT is straightforward, it still requires
hundreds of cycles (or a few microseconds) to process each log entry, which is
a significant amount in the face of sub 10 us network latency between nodes with
DPDK. We amortize this overhead when operating under load by
batching multiple RPC calls together and treating the batch as a single RAFT log
entry.  DPDK by itself can efficiently batch incoming packets under load,
therefore allowing the user to easily receive a burst of packets
(currently limited to 32) rather than a single packet for a polling call.
DPDK further allows chaining packet
buffers together and treating the chain as a single packet. This is convenient
for our zero copy approach and we chain packet mbufs together to form a single
packet. Since the entire chain is sent as a single packet when being replicated,
the length of a chain is limited by the size of the coalesced packet, that must
fit into an ethernet jumbo frame of approximately 9KB.

@@

\subsection{Distributed Transactions}
A useful primitive for many distributed systems is the ability to run
distributed transactions - an atomic sequence of conditional updates spanning
multiple groups of replicas. Distributed transactions have two distinct facets:
atomicity and fault tolerance. Distributed transactions appear atomic with
respect to other atomic transactions and simple updates or reads by virtue of
standard concurrency control techniques. A common technique is strict two phase
locking: acquire all locks, update the locked objects and finally release all
the locks. We believe that NVM application programmers - particularly those
having experience with concurrent code should not find it hard to deal with
distributed concurrency. The hard part for most programmers is likely to be the
second aspect: fault tolerance. A distributed transaction spans many machines
and has a large footprint to clean up in the event of failure. Failure handling
is made more complex when considering cascading failures - further failures
while recovering from a failure. Cyclone was designed to allow primitives such
as distributed transctions to be added without the programmer needing to worry
about faults or write fault recovery code. As an exercise we show how to add
support for distributed transactions to our hash table.

@@

We guard against faults by executing the RPC call on a quorum of machines via
Cyclone. Since execution of the RPC call leaves \emph{externally visible side
  effects} by virtue of the RPC calls it makes we require output
commit~\cite{output_commit}. This is achieved if the execution is guaranteed to
complete once it starts. This is simply achieved by a flag in the RPC call that
turns off decoupled execution (Section~\ref{sec:decouple}). The flag causes
execution to start only after the RPC call has been successfully replicated.

@@

We note that this design is not intended for performance, under the assumption
that distributed transactions will be infrequent compared to single key
operations.  On the other hand, it is worth comparing the lack of any need for
explicit fault recovery in Cyclone with fault recovery in systems systems that
explicity require such cleanup (~\cite{farm}, pp 59--63).  We believe that
such designs, which make availability easy for NVM applications are key to
increasing the adoption rate of NVM by data center operators.

@@

\section{Evaluation}
We evaluate Cyclone against its design goals by using it for up to 3-way
replication in a cluster of twelve 48 core Xeon servers. Three servers run
Cyclone's server side code, while the remaining machines run multithreaded
clients to load the system for our studies. The server side machines are each
equipped with four commodity 82599 10 Gbit ethernet controllers, while each
client machine has one 82599 ethernet controller. All the machines are connected
through the same commodity 10Gbit switch. We use 8 cores on the server machines
for running 8 instance of RAFT, leaving 39 cores to run application code. The
last remaining core is reserved for management code. We use NVML with the non
volatile pool held in DRAM, which proxies for the performance we would expect
from NVDIMMs. We evaluate Cyclone in three stages.

First, we evaluate pure replication performance by replicating RPC calls. There
is crash consistent transaction that remembers the RPC call sequence number -
every RPC call is accepted and once replicated a response is sent to the
client. Figure~\ref{fig:pure_rep} shows the results. We consider two cases, one
where the clients send an RPC call with no payload and one where they add a 512
byte dummy payload. For the no payload case the bottleneck is the message
handling rate. We note that Cyclone can handle small message replication rates
as high as \emph{8 million messages a second}. In the case of RPC calls with 512
byte payloads, the bottleneck is the available network bandwidth. We plot
transmit throughput derived from the message rate as the rate at which data is
sent to replicas from all the 8 quorum leaders and back to the clients.
It is interesting to note that with 3 replicas Cyclone achieves \emph{more than
  40 Gbps} of bandwidth. The reason for this is that the quorum leaders can end
up residing on different machines. Although replication traffic is still limited
by the receive bandwidth the responses to the clients, originating from
different machines have access to independent 4x10GigE NICs, allowing total
transmit throughput to exceed 40 Gbps. Although not discussed in this paper,
Cyclone's control plane is designed to exploit this by trying to spread quorum
leaders across machines.

Also worth noting is that thanks to DPDK's low latency access to the NICs, three
way replication costs as little as 25 us of latency in the unloaded case. This
latency is measured at the client and therefore is four commodity 10 GigE
network hops through the switch, with each hop taking under 7 us including
network stack delays. These results should serve to illustrate that simple
commodity ethernet can indeed provide reasonable replication performance with a
carefully designed software stack that can exploit it.

\begin{figure}
\begin{tabular}{cc}
\begin{minipage}{0.25\textwidth}
  \includegraphics[width=\textwidth,height=3cm]{results/replication_mops_0.pdf}
\end{minipage}&
\begin{minipage}{0.25\textwidth}
  \hspace{-0.2in}
  \includegraphics[width=\textwidth,height=3cm]{results/replication_gbps_0.pdf}
\end{minipage}\\
\begin{minipage}{0.25\textwidth}
  \includegraphics[width=\textwidth,height=3cm]{results/replication_mops_512.pdf}
\end{minipage}&
\begin{minipage}{0.25\textwidth}
  \hspace{-0.2in}
  \includegraphics[width=\textwidth,height=3cm]{results/replication_gbps_512.pdf}
\end{minipage}
\end{tabular}
\caption{Pure Replication}
\label{fig:pure_rep}
\end{figure}

Next, we consider the effect of adding NVMLs crash consistent
transactions. Figure~\ref{fig:cc_rep} shows the effect of adding a crash
consistent transaction where the application still does nothing but Cyclone
remembers the last sequence number and RPC call result (the application simply
returns the received payload). Throughput is now bottlenecked by NVML allocator
performance at just under a million messages a second. The final result is
adding the hash table in the picture with the clients updating the values of
keys already present in the hash table - a stable workload for easy
comparison. As Figure~\ref{fig:app_rep} shows performance stays at the just
under 1M calls a second mark for pure updates. For comparison we also show a
pure read only case where the clients only query the value of a key - there is
no replication \emph{nor} any crash consistent transaction applied. It is
interesting to note that the pure read only case is bottlenecked at the same
point as the pure replication case of 8 million operations a second.

These results lead us to conclude that Cyclone's provides adequate performance
headroom for replication to NVML applications. In fact our current focus in the
light of these results is towards improving application performance on top of
NVML. NVML is a general purpose library written for a variety of substrates
including traditional SSD based storage and there is adequate room for
improvement of performance for applications running on directly attached
persistent memory with close to DRAM performance.

\begin{figure}
\begin{tabular}{cc}
\begin{minipage}{0.25\textwidth}
  \includegraphics[width=\textwidth,height=3cm]{results/cc_mops.pdf}
  \caption{Crash Consistency + Replication}
  \label{fig:cc_rep}
\end{minipage} &
\begin{minipage}{0.25\textwidth}
  \includegraphics[width=\textwidth,height=3cm]{results/app_mops.pdf}
  \caption{Crash Consistent Replicated Hash Table}
  \label{fig:app_rep}
\end{minipage}
\end{tabular}
\end{figure}

\section{Conclusion}
Cyclone allows developers to add availability to applications built on top of
libraries such as NVML and achieve good performance on commodity networking
hardware. A key question is whether the ideas in Cyclone extend to more than
just key-value maps - perhaps to application such as filesystems ? We believe
that the requirement for scaling across multiple replication quorums can be
satisfied by applications where the scalable commutativity
rule~\cite{scalable_commutativity} applies to common case primitives. Although
our intitial release of Cyclone will be targeted towards key value style
applications, our future research is targeted towards investigating this
hypothesis.
\newcommand\myurl[2]{\url{#1}}
\bibliographystyle{acm}
\bibliography{paper}
